---
title: "KNN"
author: "Ballesteros"
date: "2023-03-20"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Reading Libraries
This block loads the necessary libraries for the KNN analysis and reads the "tele_norm" dataset.
```{r}
library(class)
library(gmodels)
library(caret)
```

### Cleaning Data

This block creates the training and test sets for classification using the KNN algorithm. It selects a random subset of rows as the test set and uses the remaining rows as the training set.
```{r}
tele_norm <-readRDS("tele_norm")

# Selects 10000 random rows for test data
set.seed(12345)
test_set <- sample(1:nrow(tele_norm), nrow(tele_norm)/2) 
# Depending on R-version and computer, different rows may be selected. 
# If that happens, results are different. 

# Create a train set and test set
#First the predictors - all columns except the yyes column
tele_train <- tele_norm[-test_set, -match("yyes",names(tele_norm))]
tele_test  <- tele_norm[test_set, -match("yyes",names(tele_norm))]

#Now the response (aka Labels) - only the yyes column
tele_train_labels <- tele_norm[-test_set, "yyes"]
tele_test_labels  <- tele_norm[test_set, "yyes"]

```

### Finding optimal K
This block uses a loop to calculate the Kappa statistic for a range of K values and saves the results in a data frame. The data frame is then saved to a file and a plot of Kappa values against K values is produced.
```{r eval = FALSE}
knn_df<- data.frame()
  
for (i in seq(3,145,2)){
 knn_pred <- knn(tele_train, tele_test, cl = tele_train_labels , k = i)
 CM <- confusionMatrix(as.factor(knn_pred),as.factor(tele_test_labels))
 knn_df<-rbind(knn_df,c(i,CM$overall[2]))
}

names(knn_df) <- c("k","Kappa")
saveRDS(knn_df,"optimal_k")
```

### Ploting error 
This block performs KNN regression on the same dataset to identify the optimal value of K for regression. The RMSE, R-squared, and MAE statistics are saved for each value of K in a data frame, which is then saved to a file.
```{r}
knn_df<-readRDS("optimal_k")

plot(knn_df$k, knn_df$Kappa, type = "l")

```

### Saving KNN prediction
```{r eval = FALSE}
knn_pred <- knn(tele_train, tele_test, cl = tele_train_labels , k = knn_df$k[which.max(knn_df$Kappa)])

saveRDS(knn_pred,"knn_pred")
```

This block produces three plots of the RMSE, R-squared, and MAE statistics against K values.
```{r eval = FALSE}
tele_train$yyes <- tele_train_labels

response_df<- data.frame("RMSE", "Rsquared", "MAE", "k")
for (i in 1:30){
  knn_reg_pred <- 
  knnreg(yyes~., data = tele_train, k = i)

  pred1 <- predict(knn_reg_pred, tele_test)

  response_df[i,] <-c(postResample(pred1,tele_test_labels),i)
}

saveRDS(response_df,"optimal_k_reg")
```

```{r}
response_df<-readRDS("optimal_k_reg")
plot(response_df$X.k.,response_df$X.RMSE., type = "l")
```
```{r}
plot(response_df$X.k.,response_df$X.Rsquared., type = "l")
```
```{r}
plot(response_df$X.k.,response_df$X.MAE., type = "l")
```

This block performs KNN regression with the optimal K value and saves the predicted values for the test data. It then produces a confusion matrix by comparing the predicted values to the actual test data.
```{r eval=FALSE}
knn_reg <- knnreg(tele_train_labels~., data = tele_train, k = 15 )

knn_reg_pred <- predict(knn_reg,tele_test)

saveRDS(knn_reg_pred, "knn_reg_pred")
```

```{r}
knn_reg_pred <- readRDS("knn_reg_pred")
confusionMatrix(as.factor(ifelse(knn_reg_pred>0.16,1,0)),as.factor(tele_test_labels))
```


